Reproducible Research: Peer Assignment #1
========================================================

Load the data:
---------------

```{r}
if (!file.exists("activity.csv")) {
    unzip("activity.zip")
}
activity<-read.csv("activity.csv", stringsAsFactors = FALSE)
str(activity)
```

What is the total number of steps taken per day?
----------------------------------------------------
```{r}
totsteps<- tapply(activity$steps, activity$date, sum, na.rm=TRUE)
```

What are the mean and median steps taken?
----------------------------------------------------
```{r}
mean(totsteps)
median(totsteps)
```

Make a histogram of the steps taken:
----------------------------------------------------
```{r}
require("ggplot2")
qplot(totsteps, xlab = "Total steps", ylab = "Frequency")
```

What is the average daily activity pattern?
----------------------------------------------------
```{r}
meansteps<- tapply(activity$steps, activity$interval, mean, na.rm=TRUE)
```

Plot the daily average activity pattern:
--
```{r}
plot(meansteps, type="l", col="blue", xlab="time of day", ylab="mean steps")
```

Imputing missing values:
----------------------------------------------------
Lets determine how many step values are missing
--
```{r}
summary(activity$steps)
```

Lets use the Hmisc package to impute th missing step values and create a new column for this infomation:
--
```{r}
require(Hmisc)
activityimp<- activity
activityimp$steps <- with(activityimp, impute(steps, mean))
str(activityimp$steps)
```

Lets use tapply function to get the imputed total steps
--
```{r}
totstepsimp<-tapply( activityimp$steps,  activityimp$date, sum)
```

Compare previous mean and median total steps with mean and median total imputed steps data:
--
```{r}
mean(totsteps)
mean(totstepsimp)

median(totsteps)
median(totstepsimp)
```

Make a histogram of the imputed steps data:
----------------------------------------------------
```{r}
qplot(totstepsimp, xlab = "Total steps", ylab = "Frequency")
```

Are there differences in activity patterns between weekdays and weekends?
--------------------------------------------------------------------------
Create Subsets for weekends and weekdays:
First let's change the imputed date and time data into something cleaner using as.POSIXct
--

```{r}
activityimp$date.time <- as.POSIXct(paste(activityimp$date, time), format = "%Y-%m-%d %H.%M", tz = "GMT")
```


```{r}

weekend <- subset(activityimp, weekdays(activityimp$date.time) %in% c("Saturday", "Sunday"))

weekdays <- subset(activityimp, weekdays(activityimp$date.time) %in% c("Monday", "Tuesday", "Wednesday","Thursday", "Friday"))

weekendsteps<-tapply(weekend$steps, weekend$interval, mean)
weekdaysteps<-tapply(weekdays$steps, weekdays$interval, mean)
```

Plot the weekend and weekday steps:
---
```{r}
par(mfrow=c(1,2)) 
plot(weekdaysteps, type="l", ylab="Number of Steps", xlab="Interval", main="Weekday Steps")
plot(weekendsteps, type="l", ylab="Number of Steps", xlab="Interval", main="Weekend Steps")
```










